---
output: github_document
---

<!-- README.md is generated from README.Rmd — edit README.Rmd, then run
     rmarkdown::render("README.Rmd") (needs pandoc). -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.path = "man/figures/README-", out.width = "85%", fig.align = "center",
  dev = "jpeg", dev.args = list(quality = 85), dpi = 96,
  message = FALSE, warning = FALSE
)
# render a magick image straight to the graphics device (for figures)
show_img <- function(im) {
  inf <- magick::image_info(im)
  op <- graphics::par(mar = c(0, 0, 0, 0)); on.exit(graphics::par(op))
  graphics::plot(NA, xlim = c(0, inf$width), ylim = c(inf$height, 0),
                 asp = 1, axes = FALSE, xlab = "", ylab = "")
  graphics::rasterImage(grDevices::as.raster(im), 0, inf$height, inf$width, 0)
}
```

# scalecard

<!-- badges: start -->
[![R-CMD-check](https://github.com/daliakopoulos/scalecard/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/daliakopoulos/scalecard/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`scalecard` perspective-corrects ("rectifies") and scales photographs taken with
the [**Credit Card Photography Scale**](https://pasthorizons.com/products/credit-card-photography-scale)
(Past Horizons) calibration card. The
card's three crosshair rectification targets (on a precise 50&nbsp;&times;&nbsp;20&nbsp;mm
frame), its red/yellow/blue colour patches and its 8&nbsp;cm scale-bar squares are
detected automatically and used as ground-control points for a projective
(homography) transform. The result is a flat image at a known
pixels-per-millimetre scale, so any feature can be measured in millimetres. It is
an R port of the [online browser tool](https://agroinformatics.gr/card-rectifier/)
(OpenCV.js), built on `magick`, `imager` and `EBImage`.

## Installation

```r
# needs: magick, imager, and EBImage (Bioconductor)
# install.packages(c("magick", "imager"))
# install.packages("BiocManager"); BiocManager::install("EBImage")

# install.packages("remotes")
remotes::install_github("daliakopoulos/scalecard")
```

On macOS, `imager` links against X11 — install
[XQuartz](https://www.xquartz.org/) first if you don't already have it.

## A worked example

The package ships with a sample photo (a calibration card and two objects on a
cutting mat). Load it and run the whole pipeline in one call.

```{r load}
library(scalecard)
img <- system.file("extdata", "sample.jpg", package = "scalecard")
```

### Detection

`detect_card()` finds the card and all its control points automatically.

```{r detect}
det <- detect_card(img)
det
```

The overlay shows the detected geometry: the blue outline follows the card's true
perspective skew, the green "L" is the 50&nbsp;&times;&nbsp;20&nbsp;mm target frame, yellow crosses
the three crosshair targets, magenta dots the colour patches and cyan squares the
8&nbsp;cm bar squares.

```{r overlay, fig.width = 6, fig.height = 4.5}
plot(det, image = img)
```

### Rectification

`rectify_card()` builds a homography from all the control points and warps the
photo flat at a chosen resolution (default 10&nbsp;px/mm). It white-balances off the
card, keeps the full scene (never cropped), and fills anything outside the
original frame with white instead of smeared edges.

```{r rectify}
rec <- rectify_card(img, grid = FALSE)
rec
```

```{r rectified, fig.width = 6, fig.height = 5.5}
show_img(rec$image)
```

### Measuring

The output has a known scale, so a pixel distance converts straight to
millimetres:

```{r measure}
rec$px_per_mm            # pixels per mm
measure_mm(rec, 250)     # 250 px on the rectified image, in mm
```

## It rectifies ONE plane

The card fixes a single plane (the surface it lies on). **Flat** objects on that
surface rectify to their true size and shape anywhere in the frame. Objects with
**height** (3-D shapes) distort by parallax, which a single photo cannot correct.
The farther from the card and the steeper the camera angle, the more the
surrounding plane is stretched when flattened.

**For best results:** lay the card flat *in the same plane* as the objects and
close to them, keep objects flat, and shoot as square-on (top-down) as you can
with the whole working area in frame.

## When auto-detection fails

For difficult photos (poor lighting, busy background, tiny card) use the
interactive fallback `manual_targets("photo.jpg")`: a native window opens and you
click the three crosshair targets; a matched filter snaps each click to the ring
centre. The result is passed to `rectify_card()` via its `detection` argument.

## Function reference

| function | purpose |
|---|---|
| `rectify_card()`   | detect + rectify + (optional) colour fix + grid |
| `detect_card()`    | automatic detection of all ground-control points |
| `manual_targets()` | interactive click-based fallback for the crosshairs |
| `white_balance()`  | gray-world white balance off the card |
| `color_correct()`  | colour transform to the card's reference swatches |
| `measure_mm()`     | convert a pixel distance to mm |
| `card_spec()`      | the fixed card geometry (edit if your card differs) |

A fuller walkthrough is in the package vignette: `vignette("scalecard")`.

## License

MIT &copy; Ioannis Daliakopoulos
