---
title: "Rectifying and scaling photos with the Credit Card Photography Scale"
author: "Ioannis Daliakopoulos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rectifying and scaling photos with the Credit Card Photography Scale}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  warning = FALSE, message = FALSE,
  fig.align = "center", out.width = "85%",
  dev = "jpeg", dev.args = list(quality = 85), dpi = 96
)
# render a magick image straight to the graphics device (for figures)
show_img <- function(im) {
  inf <- magick::image_info(im)
  op <- graphics::par(mar = c(0, 0, 0, 0)); on.exit(graphics::par(op))
  graphics::plot(NA, xlim = c(0, inf$width), ylim = c(inf$height, 0),
                 asp = 1, axes = FALSE, xlab = "", ylab = "")
  graphics::rasterImage(grDevices::as.raster(im), 0, inf$height, inf$width, 0)
}
```

## What it does

`scalecard` perspective-corrects ("rectifies") and scales photographs taken with
the *Credit Card Photography Scale* (Past Horizons) calibration card. The card's
three crosshair rectification targets (on a precise 50 &times; 20 mm frame), its
red/yellow/blue colour patches and its 8 cm scale-bar squares are detected
automatically and used as ground-control points for a projective (homography)
transform. The result is a flat image at a known pixels-per-millimetre scale, so
any feature can be measured in millimetres. It is an R port of a browser
(OpenCV.js) tool, built on `magick`, `imager` and `EBImage`.

## A worked example

The package ships with a sample photo (a calibration card and two objects on a
cutting mat). We load it and run the whole pipeline in one call.

```{r load}
library(scalecard)
img <- system.file("extdata", "sample.jpg", package = "scalecard")
```

### Detection

`detect_card()` finds the card and all its control points automatically.

```{r detect}
det <- detect_card(img)
det
```

We can overlay the detected geometry on the photo. The blue outline follows the
card's true perspective skew, the green "L" is the 50 &times; 20 mm target frame,
the yellow crosses are the three crosshair targets, magenta dots the colour
patches and cyan squares the 8 cm bar squares.

```{r overlay, fig.width = 6, fig.height = 4.5}
plot(det, image = img)
```

### Rectification

`rectify_card()` builds a homography from all the control points and warps the
photo flat at a chosen resolution (default 10 px/mm). It white-balances off the
card, keeps the full scene (never cropped), and fills anything outside the
original frame with white instead of smeared edges.

```{r rectify}
rec <- rectify_card(img, grid = FALSE)
rec
```

```{r show-rect, fig.width = 6, fig.height = 5.5}
show_img(rec$image)
```

### Measuring

The output has a known scale, so a pixel distance converts straight to
millimetres:

```{r measure}
rec$px_per_mm            # pixels per mm
measure_mm(rec, 250)     # 250 px on the rectified image, in mm
```

A copy with a 10 mm / 50 mm grid is handy for checking the scale against the
card's 8 cm bar:

```{r grid, fig.width = 6, fig.height = 5.5}
recg <- rectify_card(img, grid = TRUE)
show_img(recg$grid)
```

## It rectifies ONE plane

The card fixes a single plane (the surface it lies on). **Flat** objects on that
surface rectify to their true size and shape anywhere in the frame. Objects with
**height** (3-D shapes) distort by parallax, which a single photo cannot correct.
The farther from the card and the steeper the camera angle, the more the
surrounding plane is stretched when flattened.

**For best results:** lay the card flat *in the same plane* as the objects and
close to them, keep objects flat, and shoot as square-on (top-down) as you can
with the whole working area in frame.

## When auto-detection fails

For difficult photos (poor lighting, busy background, tiny card) use the
interactive fallback `manual_targets("photo.jpg")`: a native window opens and you
click the three crosshair targets; a matched filter snaps each click to the ring
centre. The result is passed to `rectify_card()` via its `detection` argument.
(This needs an interactive session and is not run here.)

## Function reference

| function | purpose |
|---|---|
| `rectify_card()`   | detect + rectify + (optional) colour fix + grid |
| `detect_card()`    | automatic detection of all ground-control points |
| `manual_targets()` | interactive click-based fallback for the crosshairs |
| `white_balance()`  | gray-world white balance off the card |
| `color_correct()`  | colour transform to the card's reference swatches |
| `measure_mm()`     | convert a pixel distance to mm |
| `card_spec()`      | the fixed card geometry (edit if your card differs) |
